Stack Safety is a Security Property

### 2021-04-13 tex/main.tex
More on 3

@@

%From the classic stack smashing attack to return oriented programming, 
One of the most pernicious ways to attack the stack is by hijacking control flow.
Our stack safety properties do not assume that control flow is well-bracketed; rather,
they express what it means to protect data even in the presence of control-flow attacks.
Consider example \ref{fig:controlflow}, a relative of the classic ``awkward example,''~\citep{??}
in which {\tt f} is called twice, at lines 2 and 4. It should therefore return to lines
3 and 5, respectively. But the first time, before returning, it stashes its
return pointer away, and the second time it returns to line 3 instead of 5; then it ends
up being called a third time, at which point it returns to line 5.
\apt{I added \# marks to the line numbers to avoid confusion; this should
  probably be explained. What do the colors mean in this example?}

@@

Passing on the stack is thus straightforward. The same mechanism can be used to pass
variables by reference, and here there are important stack safety concerns,
as seen in Figure~\ref{fig:passing}. In this example,
{\tt main} calls {\tt f} with a reference to {\tt x}, but then clears {\tt x} and instead
passes a reference to {\tt y}. We would expect that the first call to {\tt f} might modify
{\tt x}, but not the second. Unfortunately {\tt f} has hidden away our reference and modifies
{\tt x} in the second call.

@@

When passing by reference, it is possible to repeatedly pass the same reference further
and further down the stack. Note, however, that this must be a continuous chain; it will
not skip over any calls in a chain. So we continue to track two depths, for each passed
cell: the original owner, and the highest depth it has been passed to. It is considered
accessible anywhere between the two.

We do not model permanent sharing, in which a variable's address is taken and it might be
passed far afield before use, in detail. The appropriate way to do so in our model is
for such a variable to have no protections except for temporal safety -- it would be
unprotected until its original owner returns and it is deallocated. Then we would expect
to separately require memory safety of such data. This is
a separate property that is often enforced using capabilities, whether tag-based or
otherwise. In fact, memory safety is the interesting part of this scenario, but is out of
scope.

>>>>

\apt{How about an example incorporating temporal safety violations due to
   reading uninitialized values?}

@@

%From the classic stack smashing attack to return oriented programming, 
One of the most pernicious ways to attack the stack is by hijacking control flow.
Our stack safety properties do not assume that control flow is well-bracketed; rather,
they express what it means to protect data even in the presence of control-flow attacks.
Consider example \ref{fig:controlflow}, a relative of the classic ``awkward example,''~\citep{??}.
(Here we show return addresses explicitly as line numbers, in the form $\#n$.)
\apt{what do the colors mean in this example?}
Function {\tt f} is called twice, at lines 2 and 4; it should therefore return to lines
3 and 5, respectively. But the first time, before returning, it stashes its
return pointer away, and the second time it returns to line 3 instead of 5; then it ends
up being called a third time, at which point it returns to line 5.

@@

The same annotation mechanism can be used to pass
variables by reference, and here there are important stack safety concerns,
as seen in Figure~\ref{fig:passing}. In this example, where we revert to assuming
that arguments are passed in registers,
{\tt main} calls {\tt f} with a reference to {\tt x}, but then clears {\tt x} and instead
passes a reference to {\tt y}. We would expect that the first call to {\tt f} might modify
{\tt x}, but not the second. Unfortunately {\tt f} has hidden away our reference and modifies
{\tt x} in the second call.

@@

When passing by reference, it is possible to repeatedly pass the same reference further
and further down the stack. Note, however, that this sharing of the reference must be over
an unbroken chain of calls. So it suffices to track two depths for each passed
cell: the original owner, and the highest depth it has been passed to. It is considered
accessible anywhere between the two.

Our model provides only limited support for handling fully first-class pointers into
stack frames, e.g., pointers obtained by taking the address of a local variable.
Unlike the case for pass-by-reference, there is no easy way to model the intended sharing behavior
of such pointers in terms of stack frame activation patterns.  The best approximation we can make is to
treat a location whose address is taken as unsealed for the rest of its lifetime, i.e., until
its original owner deallocates it (typically by returning).  This prevents violations
of temporal safety (use after free), but still permits callees to access the location without having
been explicitly given a pointer to it.  Preventing the latter kind of abuse
is the provence of a fine-grained memory safety policy, e.g. based on capabilities
(implemented by tags or some other mechanism); we believe it is beyond the scope of stack safety
\emph{per se}.
